\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-02-06 2-the-cubble-pkg.Rmd
fix marked red sentences labelled for myseyself

\textcolor{red}{To use functionalities from cubble, data analysts first need to create a cubble.} `as_cubble()` creates a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude. The naming of `key` and `index` follows the convention in the \pkg{tsibble} package. The cubble created by default is in the nested form. Below is an example of creating a cubble: \newline

```{r echo = TRUE}
cubble_nested <- cubble::climate_flat %>%
  as_cubble(key = id, index = date, coords = c("long", "lat"))
cubble_nested
```

There are a few information in the \pkg{cubble} header: the name of the `key` variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the `ts` column with its type (`date [date], prcp [dbl], tmax [dbl], tmin [dbl]`).

@@

Notice here that the third line in the header now shows the name and type of spatial variables (`lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]`).

@@

\textcolor{red}{As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.} Function `migrate()` moves the spatial variables from the `spatial` attribute into the long cubble:

```{r}
cubble_long %>% migrate(long, lat)
```

<!-- In the workflow described above, data objects come into cubble in the nested form, then various operations on the spatial and temporal dimension can go back and forth between the nested and long form, and finally, the data will come out of cubble in the long form for further modelling or visualisation. -->

Building from an underlying `tbl_df` structure, it is natural to implement methods available in `dplyr` to `cubble`. Supported methods in the `cubble` with `dplyr` generics includes:

```{=tex}
\begin{center}
\begin{tabular}{ | m{5em} | m{15cm}| } 
basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
grouping &  \textbf{group\_by}, \textbf{ungroup}\\
slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
\end{tabular}
\end{center}
```

\pkg{cubble} is also compatible with \pkg{tsibble} where each element list-column in the nested form can be a `tbl_ts` object. \textcolor{red}{Duplicates and gaps should be first checked before structuring the data into a cubble.} If the input data is a \pkg{tsibble} object, the long form cubble is also a \pkg{tsibble} where users can directly apply time series operations. \textcolor{red}{more experiment with tsibble}

>>>>

`as_cubble()` creates a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude. The naming of `key` and `index` follows the convention in the \pkg{tsibble} package. The cubble created by default is in the nested form. Below is an example of creating a cubble:

```{r echo = TRUE}
cubble_nested <- cubble::climate_flat %>%
  as_cubble(key = id, index = date, coords = c(long, lat))
cubble_nested
```

There are a few information in the \pkg{cubble} header: the name of the `key` variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the `ts` column with its type: `date [date], prcp [dbl], tmax [dbl], tmin [dbl]`.

@@

Notice here that the third line in the header now shows the name and type of spatial variables: `lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]`.

@@

Some functions require time invariant variables to be in the same table as the time varying ones. In cubble, this can be done through `migrate()`, which moves the time invariant variables into the long cubble:

```{r echo = TRUE}
cubble_mig <- cubble_long %>% migrate(long, lat)
cubble_mig
```

Notice that these added time invariant variables will not be remembered the cubble and will disappear if switched to the nested form and then switched back. Hence they should most likely been used at the last step of the analysis:

```{r echo = TRUE}
cubble_mig %>% tamp() %>% stretch()
```

## Compatibility with existing packages

Building from an `tbl_df` class, `cubble` has implemented methods for `dplyr` generics, which includes:

  - basic: mutate, filter, summarise, select, arrange, rename, left_join
  - grouping: group_by, ungroup
  - slice family: slice_head, slice_tail, slice_sample, slice_min, slice_max

\pkg{cubble} is also compatible with \pkg{tsibble}. When creating a cubble from a tsibble, Only the `coords` argument need to be specified:

```{r echo = TRUE}
# casting a tsibble into cubble
dt <- climate_flat %>% 
  tsibble::as_tsibble(key = id, index = date) %>% 
  cubble::as_cubble(coords = c(long, lat))
dt
```

In the created nested form, each element of the list-column `ts` is of `tbl_ts` class so that existing operations on the tsibble class can still be applied:

```{r echo = TRUE}
# add station-based features in the nested form.
dt %>% 
  mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
```

### 2022-02-06 3-advanced.Rmd
fix marked red sentences labelled for myseyself

Spatial locations can have grouping structure that is either inherent to the data  (i.e. countries are nested in continents), or formed by clustering. Rather than analysing variables in the site level, summarised variables in the cluster level can give a clearer picture of \textcolor{red}{local dynamic}. In cubble, `switch_key()` can be used to create a new level of grouping of spatial location by specifying a clustering variable. Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at site and cluster level, in both the long and nested form. \textcolor{red}{unify the use of site/ station}. By specifying `cluster_nested <- station_nested %>% switch_key(key = cluster)`, the cubble re-defines the cubble key from `id` in `station_nested` to `cluster` in `cluster_nested`. All the spatial variables variant to `cluster` are now nested into a `.val` column and cluster level variables can be computed in the same fashion as station level variables in `station_nested`. Following the same principal of `stretch`, the long form cluster level cubble expands the time series variables with both indices: `cluster` and `id`. All the cluster level variables are stored separately for recovering the nested from from the long form.

>>>>

Spatial locations can have grouping structure that is either inherent to the data  (i.e. countries are nested in continents), or formed by clustering. Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas. In cubble, `switch_key()` can be used to create a new level of grouping of spatial location by specifying a clustering variable. Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form. By specifying `cluster_nested <- station_nested %>% switch_key(key = cluster)`, the cubble re-defines the cubble key from `id` in `station_nested` to `cluster` in `cluster_nested`. All the spatial variables variant to `cluster` are now nested into a `.val` column and cluster level variables can be computed in the same fashion as station level variables in `station_nested`. Following the same principal of `stretch`, the long form cluster level cubble expands the time series variables with both indices: `cluster` and `id`. All the cluster level variables are stored separately for recovering the nested from from the long form.

### 2022-02-06 4-1-basic.Rmd
fix marked red sentences labelled for myseyself

Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations. All the stations start recording from year `r range(weatherdata::historical_tmax$first_year)[2]` with the longest back to year `r range(weatherdata::historical_tmax$first_year)[1]`. The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c("longitude", "latitude")` as the coordinates. This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.

Stations in the two states can be subsetted on the station number: Australia GHCN station number starts with "ASN00" and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) define the state of the station. New South Wales stations start from 46 to 75 and Victoria stations follow from 76 to 90. Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:

@@

Filtering for the period 1971: 1975 and 2016:2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by `stretch()`:

@@

```{r eval=  FALSE, echo = TRUE}
tmax <- tmax %>%
  mutate(month = lubridate::month(date), 
         group = as.factor(ifelse(lubridate::year(date) > 2015, 
                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
  group_by(month, group) %>%
  summarise(tmax = mean(tmax, na.rm = TRUE))
```

>>>>

Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations. All the stations start recording from year `r range(weatherdata::historical_tmax$first_year)[2]` with the longest back to year `r range(weatherdata::historical_tmax$first_year)[1]`. The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c(longitude, latitude)` as the coordinates. This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.

Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with "ASN00" and followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml), where the 2nd and 3rd digit (7th and 8th in the GHCN number) define the state of the station. New South Wales stations start from 46 to 75 and Victoria stations follow from 76 to 90. Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:

@@

Filtering for the period 1971 ~ 1975 and 2016 ~ 2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by `stretch()`:

@@

```{r eval=  FALSE, echo = TRUE}
tmax <- tmax %>%
  group_by(month = lubridate::month(date), 
         group = as.factor(ifelse(lubridate::year(date) > 2015, 
                                  "2016 ~ 2020", "1971 ~ 1975"))) %>%
  summarise(tmax = mean(tmax, na.rm = TRUE))
```

### 2022-02-06 main.tex
fix marked red sentences labelled for myseyself

\textcolor{red}{To use functionalities from cubble, data analysts first need to create a cubble.}
\texttt{as\_cubble()} creates a \pkg{cubble} by supplying the three key
components: \texttt{key} as the spatial identifier; \texttt{index} as
the temporal identifier; and a vector of \texttt{coords} in the order of
longitude first and then latitude. The naming of \texttt{key} and
\texttt{index} follows the convention in the \pkg{tsibble} package. The
cubble created by default is in the nested form. Below is an example of
creating a cubble: \newline

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested <- cubble::climate_flat %>%
+   as_cubble(key = id, index = date, coords = c("long", "lat"))
R> cubble_nested
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

Notice here that the third line in the header now shows the name and
type of spatial variables
(\texttt{lat\ {[}dbl{]},\ long\ {[}dbl{]},\ elev\ {[}dbl{]},\ name\ {[}chr{]},\ wmo\_id\ {[}dbl{]}}).

@@

\textcolor{red}{As a final data output for modelling or visualisation, spatio-temporal data is usually expected to be in a single table.}
Function \texttt{migrate()} moves the spatial variables from the
\texttt{spatial} attribute into the long cubble:

\begin{CodeChunk}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin  long   lat
  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

Building from an underlying \texttt{tbl\_df} structure, it is natural to
implement methods available in \texttt{dplyr} to \texttt{cubble}.
Supported methods in the \texttt{cubble} with \texttt{dplyr} generics
includes:

\begin{center}
\begin{tabular}{ | m{5em} | m{15cm}| } 
basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
grouping &  \textbf{group\_by}, \textbf{ungroup}\\
slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
\end{tabular}
\end{center}

\pkg{cubble} is also compatible with \pkg{tsibble} where each element
list-column in the nested form can be a \texttt{tbl\_ts} object.
\textcolor{red}{Duplicates and gaps should be first checked before structuring the data into a cubble.}
If the input data is a \pkg{tsibble} object, the long form cubble is
also a \pkg{tsibble} where users can directly apply time series
operations. \textcolor{red}{more experiment with tsibble}

@@

Spatial locations can have grouping structure that is either inherent to
the data (i.e.~countries are nested in continents), or formed by
clustering. Rather than analysing variables in the site level,
summarised variables in the cluster level can give a clearer picture of
\textcolor{red}{local dynamic}. In cubble, \texttt{switch\_key()} can be
used to create a new level of grouping of spatial location by specifying
a clustering variable. Figure \ref{fig:illu-hier} illustrates the
relationship of cubbles at site and cluster level, in both the long and
nested form. \textcolor{red}{unify the use of site/ station}. By
specifying
\texttt{cluster\_nested\ \textless{}-\ station\_nested\ \%\textgreater{}\%\ switch\_key(key\ =\ cluster)},
the cubble re-defines the cubble key from \texttt{id} in
\texttt{station\_nested} to \texttt{cluster} in
\texttt{cluster\_nested}. All the spatial variables variant to
\texttt{cluster} are now nested into a \texttt{.val} column and cluster
level variables can be computed in the same fashion as station level
variables in \texttt{station\_nested}. Following the same principal of
\texttt{stretch}, the long form cluster level cubble expands the time
series variables with both indices: \texttt{cluster} and \texttt{id}.
All the cluster level variables are stored separately for recovering the
nested from from the long form.

@@

Global Historical Climatology Network (GHCN) provides daily climate
measures from stations across the world and the dataset
\texttt{weatherdata::historical\_tmax} extracts the maximum temperature
for 236 Australian stations. All the stations start recording from year
1969 with the longest back to year 1859. The data
\texttt{historical\_tmax} is already presented as a cubble, with
\texttt{id} as the key, \texttt{date} as the index, and
\texttt{c("longitude",\ "latitude")} as the coordinates. This example
compares the maximum temperature in two periods: 1971 - 1975 and 2016 -
2020 for stations in Victoria and New South Wales.

Stations in the two states can be subsetted on the station number:
Australia GHCN station number starts with ``ASN00'' and followed by the
\href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of
Meteorology (BOM) station number}, where the 2nd and 3rd digit (7th and
8th in the GHCN number) define the state of the station. New South Wales
stations start from 46 to 75 and Victoria stations follow from 76 to 90.
Filtering Victoria and New South Wales stations is an operation in the
spatial dimension and hence uses the nested form:

@@

Filtering for the period 1971: 1975 and 2016:2020 is an operation on the
time dimension and the nested cubble needs to be switched to the long
cubble by \texttt{stretch()}:

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax %>%
+   mutate(month = lubridate::month(date), 
+          group = as.factor(ifelse(lubridate::year(date) > 2015, 
+                                   "2016 ~ 2020", "1971 ~ 1975"))) %>%
+   group_by(month, group) %>%
+   summarise(tmax = mean(tmax, na.rm = TRUE))
\end{CodeInput}
\end{CodeChunk}

>>>>

\texttt{as\_cubble()} creates a \pkg{cubble} by supplying the three key
components: \texttt{key} as the spatial identifier; \texttt{index} as
the temporal identifier; and a vector of \texttt{coords} in the order of
longitude first and then latitude. The naming of \texttt{key} and
\texttt{index} follows the convention in the \pkg{tsibble} package. The
cubble created by default is in the nested form. Below is an example of
creating a cubble: \newline

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested <- cubble::climate_flat %>%
+   as_cubble(key = id, index = date, coords = c(long, lat))
R> cubble_nested
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

Notice here that the third line in the header now shows the name and
type of spatial variables:
\texttt{lat\ {[}dbl{]},\ long\ {[}dbl{]},\ elev\ {[}dbl{]},\ name\ {[}chr{]},\ wmo\_id\ {[}dbl{]}}.

@@

Some functions require time invariant variables to be in the same table
as the time varying ones. In cubble, this can be done through
\texttt{migrate()}, which moves the time invariant variables into the
long cubble:

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig <- cubble_long %>% migrate(long, lat)
R> cubble_mig
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin  long   lat
  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

Notice that these added time invariant variables will not be remembered
the cubble and will disappear if switched to the nested form and then
switched back. Hence they should most likely been used at the last step
of the analysis:

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig %>% tamp() %>% stretch()
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{compatibility-with-existing-packages}{%
\subsection{Compatibility with existing
packages}\label{compatibility-with-existing-packages}}

Building from an \texttt{tbl\_df} class, \texttt{cubble} has implemented
methods for \texttt{dplyr} generics, which includes:

\begin{itemize}
\tightlist
\item
  basic: mutate, filter, summarise, select, arrange, rename, left\_join
\item
  grouping: group\_by, ungroup
\item
  slice family: slice\_head, slice\_tail, slice\_sample, slice\_min,
  slice\_max
\end{itemize}

\pkg{cubble} is also compatible with \pkg{tsibble}. When creating a
cubble from a tsibble, Only the \texttt{coords} argument need to be
specified. In the created nested form, each element of the list-column
\texttt{ts} is of \texttt{tbl\_ts} class so that existing operations on
the tsibble class can still be applied:

\begin{CodeChunk}
\begin{CodeInput}
R> # casting a tsibble into cubble
R> dt <- climate_flat %>% 
+   tsibble::as_tsibble(key = id, index = date) %>% 
+   cubble::as_cubble(coords = c(long, lat))
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
\end{CodeOutput}
\begin{CodeInput}
R> # add station-based features in the nested form.
R> dt %>% 
+   mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts      tmax_mean tmax_var
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>      <dbl>    <dbl>
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_t~      25.7    38.6 
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_t~      26.2    51.1 
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_t~      23.7    45.4 
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_t~      33.1     3.02
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_t~      34.6    24.7 
\end{CodeOutput}
\end{CodeChunk}

@@

Spatial locations can have grouping structure that is either inherent to
the data (i.e.~countries are nested in continents), or formed by
clustering. Rather than analysing variables in the site level,
summarised variables in the cluster level can give a crisper picture of
local areas. In cubble, \texttt{switch\_key()} can be used to create a
new level of grouping of spatial location by specifying a clustering
variable. Figure \ref{fig:illu-hier} illustrates the relationship of
cubbles at station and cluster level, in both the long and nested form.
By specifying
\texttt{cluster\_nested\ \textless{}-\ station\_nested\ \%\textgreater{}\%\ switch\_key(key\ =\ cluster)},
the cubble re-defines the cubble key from \texttt{id} in
\texttt{station\_nested} to \texttt{cluster} in
\texttt{cluster\_nested}. All the spatial variables variant to
\texttt{cluster} are now nested into a \texttt{.val} column and cluster
level variables can be computed in the same fashion as station level
variables in \texttt{station\_nested}. Following the same principal of
\texttt{stretch}, the long form cluster level cubble expands the time
series variables with both indices: \texttt{cluster} and \texttt{id}.
All the cluster level variables are stored separately for recovering the
nested from from the long form.

@@

Global Historical Climatology Network (GHCN) provides daily climate
measures from stations across the world and the dataset
\texttt{weatherdata::historical\_tmax} extracts the maximum temperature
for 236 Australian stations. All the stations start recording from year
1969 with the longest back to year 1859. The data
\texttt{historical\_tmax} is already presented as a cubble, with
\texttt{id} as the key, \texttt{date} as the index, and
\texttt{c(longitude,\ latitude)} as the coordinates. This example
compares the maximum temperature in two periods: 1971 - 1975 and 2016 -
2020 for stations in Victoria and New South Wales.

Stations in the two states can be subsetted by the station number:
Australia GHCN station number starts with ``ASN00'' and followed by the
\href{http://www.bom.gov.au/climate/cdo/about/site-num.shtml}{Bureau of
Meteorology (BOM) station number}, where the 2nd and 3rd digit (7th and
8th in the GHCN number) define the state of the station. New South Wales
stations start from 46 to 75 and Victoria stations follow from 76 to 90.
Filtering Victoria and New South Wales stations is an operation in the
spatial dimension and hence uses the nested form:

@@

Filtering for the period 1971 \textasciitilde{} 1975 and 2016
\textasciitilde{} 2020 is an operation on the time dimension and the
nested cubble needs to be switched to the long cubble by
\texttt{stretch()}:

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax %>%
+   group_by(month = lubridate::month(date), 
+          group = as.factor(ifelse(lubridate::year(date) > 2015, 
+                                   "2016 ~ 2020", "1971 ~ 1975"))) %>%
+   summarise(tmax = mean(tmax, na.rm = TRUE))
\end{CodeInput}
\end{CodeChunk}
